---
title: "project.rmd"
output: word_document
date: "2023-03-27"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE)

#Read Neccessary Libraries

library(datasets)
library(dplyr)
library(tidyverse)
library(lubridate)
library(plyr)
library(ggplot2)
library(purrr)
library(treemapify)
library(DBI)
library(RSQLite)
library(tidyverse)
library(igraph)

library(ggraph)
library(tidyr)
library(dplyr)
library(tidyr)
library(purrr)
library(zoo)
library(pROC)
```


## Question 1 : When is the best time of day, day of the week, and time of year to fly to minimise delays?


```{r}


#Read The Dat
plane_data<-read.csv("D:/R files/plane-data.csv")
airports<-read.csv("D:/R files/airports.csv")

conn <- dbConnect(RSQLite::SQLite(), "Coursework_1234.db")
# read the data from the 'my_table' table into a data.frame


# Drop table if it exists
dbExecute(conn, "DROP TABLE IF EXISTS d40")


# read d1 CSV file into d38 table
d1 <- read.csv("D:/R files/2004.csv")
dbWriteTable(conn, 'd39', d1, append=FALSE)

# read d2 CSV file into d38 table
d2 <- read.csv("D:/R files/2005.csv")
dbWriteTable(conn, 'd39', d2, append=TRUE)

# read d3 CSV file into d38 table
d3 <- read.csv("D:/R files/2006.csv")
dbWriteTable(conn, 'd39', d3, append=TRUE)

# remove d1, d2, d3 from memory
rm(d1, d2, d3)

 
df<- dbGetQuery(conn, "
Select * from d39 where Diverted==0 and Cancelled==0
")

dbDisconnect(conn)
```

```{r ,}
# Convert the CRSDepTime variable to a standard format (i.e. HH:MM)
dep_time_str <- sprintf("%04d", df$CRSDepTime)
dep_time_str <- substr(dep_time_str, 1, 2) %>% paste0(":", substr(dep_time_str, 3, 4))


# Combine the Year, Month, DayofMonth, and DepTime columns into a 
#single string in "YYYY-MM-DD HH:MM" format
date_str <- paste(df$Year, sprintf("%02d", df$Month), sprintf("%02d", df$DayofMonth), sep = "-")
date_str<-paste(date_str,dep_time_str,sep=' ')


# Convert the resulting string to a POSIXct date-time object
df$date <- as.POSIXct(date_str, format = "%Y-%m-%d %H:%M")

#Make new column called season based on date
df1<-df%>%
  mutate(data=make_datetime(Year,Month,DayofMonth),season=if_else
         ((data>='2003-12-21'&data<'2004-03-20')|(data>='2004-12-21'&data<'2005-03-20')|
             (data>='2005-12-21'&data<'2006-03-20')|(data>='2006-12-21'&data<'2007-03-20'),"Winter",
                                                                if_else((data>='2004-03-20'&data<'2004-06-21')|(data>='2005-03-20'&data<'2005-06-21')|
                                                                          (data>='2006-03-20'&data<'2006-06-21'),"Spring",
                                                                                  if_else((data>='2004-06-21'&data<'2004-09-23')|(data>='2005-06-21'&data<'2005-09-23')|
                                                                                            (data>='2006-06-21'&data<'2006-09-23'),"Summer",
                                                                                          if_else((data>='2004-09-23'&data<'2004-12-21')|(data>='2005-09-23'&data<'2005-12-21')|
                                                                                                    (data>='2006-09-23'&data<'2006-12-21'),"Fall","1")))))



```


```{r}
#Remove NA elmenentes from dataframe df1
df1<-df1[complete.cases(df1$ArrDelay),]

```

```{r }

#Create new column TimeofDay  using mutate and based on hour(date) which returns hour for each row
df1<-df1  %>%
   mutate(TimeofDay = case_when(
    hour(date) > 5 & hour(date) <= 11 ~ 'Morning',
    hour(date) > 11 & hour(date) <= 17 ~ 'Noon',
    hour(date) > 17 & hour(date) <= 24 ~ 'Night',
    hour(date) >= 0 & hour(date) <= 5 ~ 'Midnight'
  ))

#Create new column Group_Hour  using mutate and based on hour(date) which returns hour for each row
df1<-df1  %>%
   mutate(Group_Hour = case_when(
    hour(date) >= 0 & hour(date) <= 2 ~ '0-2 AM',
    hour(date) > 2 & hour(date) <= 4 ~ '2-4 AM',
    hour(date) > 4 & hour(date) <= 6 ~ '4-6 AM',
    hour(date) > 6 & hour(date) <= 8 ~ '6-8 AM',
    hour(date) > 8 & hour(date) <= 10 ~ '8-10 AM',
    hour(date) > 10 & hour(date) <= 12 ~ '10-12 AM',
    hour(date) > 12 & hour(date) <= 14 ~ '12-14 PM',
    hour(date) > 14 & hour(date) <= 16 ~ '14-16 PM',
    hour(date) > 16 & hour(date) <= 18 ~ '16-18 PM',
    hour(date) > 18 & hour(date) <= 20 ~ '18-20 PM',
    hour(date) > 20 & hour(date) <= 22 ~ '20-22 PM',
    hour(date) > 22 & hour(date) <= 24 ~ '22-24 PM'
  ))




```



# Best season to travel to minimise Delays ?

```{r }

#Filter out rows where the "ArrDelay" column is NA (not available).
#Group the data by season.
#Calculate the mean "ArrDelay" for each season.
#Join this result with a summary of the total number of flights for each season.
#Join this result with a summary of the number of delayed flights (where "ArrDelay" is greater than 0) for each season.
#Calculate the percentage of flights that were delayed for each season by dividing the number of delayed flights by the total number of flights.
#Store the final result in the "df_plot" variable.

df_plot<-df1%>%
  dplyr::filter(!is.na(df1$ArrDelay))%>%
  dplyr::group_by(season)%>%
  dplyr::summarise(mean_delay=mean(ArrDelay))%>%
  inner_join(df1%>%dplyr::group_by(season)%>%
               dplyr::summarise(total_flights=n()),by="season")%>%
  inner_join(df1%>%dplyr::filter(!is.na(df1$ArrDelay)&df1$ArrDelay>0)%>%dplyr::group_by(season)%>%
               dplyr::summarise(cnt_delay=n()),by="season")%>%
  mutate(pct_delay=cnt_delay / total_flights)

# Convert season variable to factor
df_plot$season <- factor(df_plot$season, levels = c("Winter", "Spring", "Summer", "Fall"))

df_plot


```

```{r }
# Sort data frame by smallest mean_delay value

df_plot_sorted <- df_plot %>% arrange(mean_delay)



# Create the bubble chart using GGPLOT
ggplot(df_plot_sorted, aes(x = season, y = mean_delay, size = mean_delay, color = pct_delay)) +
  geom_point(alpha = 0.7) +
  scale_size(range = c(1, 10)) +
  scale_color_gradient(low = "blue", high = "red") +
  labs(title = "Average Delay by Season", x = "Season", y = "Mean Delay (minutes)") +
  theme_minimal() +
  scale_x_discrete(limits = c("Winter", "Spring", "Summer", "Fall"))+theme_bw()
 
```

We observe that during Spring season we have less delays based on the percentage of flights that had delays is 48.45% and the mean time of the delays is the lowest at 5.9 minutes compared to other seasons

#Best season to travel to minimise Delays ?

```{r }
df_plot<-df1%>%
  dplyr::filter(!is.na(df1$ArrDelay))%>%
  dplyr::group_by(DayOfWeek)%>%
  dplyr::summarise(mean_delay=mean(ArrDelay))%>%
  inner_join(df1%>%dplyr::group_by(DayOfWeek)%>%
               dplyr::summarise(total_flights=n()),by="DayOfWeek")%>%
  inner_join(df1%>%dplyr::filter(!is.na(df1$ArrDelay)&df1$ArrDelay>0)%>%dplyr::group_by(DayOfWeek)%>%
               dplyr::summarise(cnt_delay=n()),by="DayOfWeek")%>%
  mutate(pct_delay=(cnt_delay / total_flights)*100)


ggplot(df1[,c('DayOfWeek','ArrDelay')], aes(x = DayOfWeek, y = ArrDelay)) +
  stat_summary(fun = "mean", geom = "line", size = 1, color = "steelblue") +
  geom_text(stat = "summary", fun = mean, aes(label = round(..y.., 1)), vjust = -0.5) +
  labs(x = "Day of Week", y = "Mean Arrival Delay") +
scale_x_continuous(breaks = 1:7,labels=c('Mon','Tue','Wed','Thurs','Fri','Sat','Sun'))+theme_bw()

ggplot(df_plot, aes(x = DayOfWeek, y = pct_delay)) +
  geom_bar(stat = "identity", fill = "steelblue") +geom_text(stat = "summary", fun = "mean", aes(label = round(..y.., 2)), vjust = -1, position = position_stack(vjust = 0.5))+
  scale_x_continuous(breaks = 1:7,labels=c('Mon','Tue','Wed','Thurs','Fri','Sat','Sun')) +
  labs(x = "Day of Week", y = "Percentage of Delayed Flights")+theme_bw()


```




# Best Month to travel to minimise Delays ?
```{r }

df_plot<-df1%>%
  dplyr::filter(!is.na(df1$ArrDelay))%>%
  dplyr::group_by(Month)%>%
  dplyr::summarise(mean_delay=mean(ArrDelay))%>%
  inner_join(df1%>%dplyr::group_by(Month)%>%
               dplyr::summarise(total_flights=n()),by="Month")%>%
  inner_join(df1%>%dplyr::filter(!is.na(df1$ArrDelay)&df1$ArrDelay>0)%>%dplyr::group_by(Month)%>%
               dplyr::summarise(cnt_delay=n()),by="Month")%>%
  mutate(pct_delay=(cnt_delay / total_flights)*100)

ggplot(df1[,c('Month','ArrDelay')], aes(x = Month, y = ArrDelay)) +
  geom_bar(stat = "summary", fun = "mean", fill = "steelblue") +
  geom_text(stat = "summary", fun = "mean", aes(label = round(..y.., 2)), vjust = -1, position = position_stack(vjust = 0.5))  +
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) +
  labs(x = "Month", y = "Mean Arrival Delay")+theme_bw()

ggplot(df_plot, aes(x = Month, y = pct_delay)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(stat = "summary", fun = "mean", aes(label = round(..y.., 2)), vjust = -1, position = position_stack(vjust = 0.5))+
  scale_x_continuous(breaks = 1:12, labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) +
  labs(x = "Month", y = "Percentage of Delayed Flights")+theme_bw()

```


#Best Time of Day to travel to minimise Delays ?
```{r }


df_plot<-df1%>%
  dplyr::filter(!is.na(df1$ArrDelay))%>%
  dplyr::group_by(TimeofDay)%>%
  dplyr::summarise(mean_delay=mean(ArrDelay))%>%
  inner_join(df1%>%dplyr::group_by(TimeofDay)%>%
               dplyr::summarise(total_flights=n()),by="TimeofDay")%>%
  inner_join(df1%>%dplyr::filter(!is.na(df1$ArrDelay)&df1$ArrDelay>0)%>%dplyr::group_by(TimeofDay)%>%
               dplyr::summarise(cnt_delay=n()),by="TimeofDay")%>%
  mutate(pct_delay=(cnt_delay / total_flights)*100)

ggplot(df_plot, aes(x = reorder(TimeofDay,pct_delay), y = pct_delay)) +
  geom_bar(stat = "identity", fill = "steelblue")  +geom_text(stat = "summary", fun = "mean", aes(label = round(..y.., 2)), vjust = -1, position = position_stack(vjust = 0.5))+
  labs(x = "Hour of Day", y = "Percentage of Delayed Flights (%)")+theme_bw()


ggplot(df_plot, aes(x = reorder(TimeofDay, total_flights), y = total_flights, fill = TimeofDay)) +
  geom_bar(stat = "identity",fill='steelblue') +
  geom_text(aes(label = total_flights), vjust = -0.5) +
  labs(x = "Group Hour", y = "Count", fill = "Group Hour") +
  theme_bw()+ggtitle('Count of flights for grouped hours ')+theme_bw()


ggplot(df_plot, aes(x = reorder(TimeofDay, mean_delay), y = mean_delay, fill = TimeofDay)) +
  geom_bar(stat = "identity",fill='steelblue') +
 geom_text(stat = "summary", fun = "mean", aes(label = round(..y.., 2)), vjust = -1, position = position_stack(vjust = 0.5)) +
  labs(x = "Group Hour", y = "Count", fill = "Group Hour") +
  theme_bw()+ggtitle('Mean of delay flights  for grouped hours ')+theme_bw()
```

# Best Time of Day grouped by every two hours  to travel to minimise Delays ?

```{r }

  dplyr::filter(!is.na(df1$ArrDelay))%>%
  dplyr::group_by(Group_Hour)%>%
  dplyr::summarise(mean_delay=mean(ArrDelay))%>%
  inner_join(df1%>%dplyr::group_by(Group_Hour)%>%
               dplyr::summarise(total_flights=n()),by="Group_Hour")%>%
  inner_join(df1%>%dplyr::filter(!is.na(df1$ArrDelay)&df1$ArrDelay>0)%>%dplyr::group_by(Group_Hour)%>%
               dplyr::summarise(cnt_delay=n()),by="Group_Hour")%>%
  mutate(pct_delay=(cnt_delay / total_flights)*100)


ggplot(df_plot, aes(x = reorder(Group_Hour,pct_delay), y = pct_delay)) +
  geom_bar(stat = "identity", fill = "steelblue")  +geom_text(stat = "summary", fun = "mean", aes(label = round(..y.., 2)), vjust = -1, position = position_stack(vjust = 0.5))+
  labs(x = "Hour of Day", y = "Percentage of Delayed Flights (%)")+theme_bw()


ggplot(df_plot, aes(x = reorder(Group_Hour, total_flights), y = total_flights, fill = Group_Hour)) +
  geom_bar(stat = "identity",fill='steelblue') +
  geom_text(aes(label = total_flights), vjust = -0.5) +
  labs(x = "Group Hour", y = "Count", fill = "Group Hour") +
  theme_bw()+ggtitle('Count of flights for grouped hours ')+theme_bw()


ggplot(df_plot, aes(x = reorder(Group_Hour, mean_delay), y = mean_delay, fill = Group_Hour)) +
  geom_bar(stat = "identity",fill='steelblue') +
 geom_text(stat = "summary", fun = "mean", aes(label = round(..y.., 2)), vjust = -1, position = position_stack(vjust = 0.5)) +
  labs(x = "Group Hour", y = "Count", fill = "Group Hour") +
  theme_bw()+ggtitle('Mean of delay flights  for grouped hours ')+theme_bw()





```




